Defficiency Screen Functional Analysis

### 2025-06-13 DeficiencyScreen/Prelim_Analysis.Rmd
fixed Df location on 2

DfKey = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/DfKey.csv"), row.names = 1)
old = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/DeficiencyModifierGenes.csv"),row.names = 1)

GeneIDKey = read.csv(paste0(git.dir,"GeneralDataFiles/GeneIDKey.csv"),row.names = 1)

dfdata = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/20250602_UpdatedModiferData.csv"))
row.names(dfdata) = dfdata$Fbab
dfgenes = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/GenesInDeficiency.csv"))

fulldf = cbind(dfgenes, dfdata[dfgenes$FBab,])
row.names(fulldf) = 1:nrow(fulldf)
fulldf$Modifier[fulldf$Modifier ==  "Supressor"] = "Suppressor"
fulldf$Modifier[fulldf$Modifier ==  "No effect"] = "No Effect"
new=fulldf

olddf = DfKey$Deficiency
newdf = unique(new$Deficiency)
oldfb = old$FBgn
newfb = unique(new$FBgn)
oldsym = old$Symbol
newsym = unique(new$Symbol)

#old = cbind(old, GeneIDKey[row.names(old), c("Start", "End", "CHR")])
old = GeneIDKey[, c("FBgn", "Start", "End", "CHR")]

old.left = apply(old[,c("Start", "End")], MARGIN = 1, min)
old.right = apply(old[,c("Start", "End")], MARGIN = 1, max)
Df.left = apply(DfKey[,c("Start", "End")], MARGIN = 1, min)
Df.right = apply(DfKey[,c("Start", "End")], MARGIN = 1, max)

genesinolddf = list()
i=1
while(i <= nrow(DfKey)){
  temp = na.omit(old[((old.left <= Df.left[i] & old.right >= Df.right[i]) |
           (old.left >= Df.left[i] & old.left <= Df.right[i]) |
           (old.right >= Df.left[i] & old.right <= Df.right[i])) &
            old$CHR == DfKey$Chr[i],])

  genesinolddf[[DfKey$Deficiency[i]]] = temp$FBgn
  i=i+1
}

#new.loc = GeneIDKey[new$FBgn,c("Start","End","CHR")]
new.loc = GeneIDKey[,c("FBgn", "Start","End","CHR")]
new.left = apply(new.loc[,c("Start", "End")], MARGIN = 1, min)
new.right = apply(new.loc[,c("Start", "End")], MARGIN = 1, max)
newDf.left = apply(new[,c("Start", "End")], MARGIN = 1, min)
newDf.right = apply(new[,c("Start", "End")], MARGIN = 1, max)

new.keep = !duplicated(paste0(new$Deficiency, newDf.left, newDf.right))
newDf.left = newDf.left[new.keep]
newDf.right = newDf.right[new.keep]
newDf = new$Deficiency[new.keep]
newChr = new$Chr[new.keep]

genesinnewdf = list()
i=1
while(i <= length(newDf)){
      temp = na.omit(new.loc[((new.left <= newDf.left[i] & new.right >= newDf.right[i]) |
           (new.left >= newDf.left[i] & new.left <= newDf.right[i]) |
           (new.right >= newDf.left[i] & new.right <= newDf.right[i])) &
            new.loc$CHR == newChr[i],])

    genesinnewdf[[newDf[i]]] = temp$FBgn
  i=i+1
}

>>>>

##DfKey = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/DfKey.csv"))
##colnames(DfKey)[c(1,3)] = c("BDSC", "FBab")
##DfKey$FollowUp = "FALSE"
##DfKey$Excluded = "FALSE"

#old.dfdata = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/DeficiencyModifierGenes.csv"),row.names)

##GeneIDKey = read.csv(paste0(git.dir,"GeneralDataFiles/GeneIDKey.csv"),row.names = 1)

##dfdata = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/20250602_UpdatedModiferData.csv"))
##colnames(dfdata)[3] = "FBab"
##dfdata$Excluded = "FALSE"

##DfKey[DfKey$FBab %in% setdiff(DfKey$FBab, dfdata$FBab), "Excluded"] = "TRUE"

##fullDfKey = rbind(dfdata, DfKey[DfKey$Excluded == TRUE, colnames(dfdata)])

##fullDfKey$Modifier[fullDfKey$Modifier == "No effect"] = "No Effect"

##fullDfKey$Modifier[fullDfKey$Modifier == "Supressor"] = "Suppressor"

##fullDfKey$Modifier[fullDfKey$Modifier == "Strong Suppressor"] = "Suppressor"

##changed start location for FBab0002465
###was listed as 1 but on Flybase it listed a different start site

##write.csv(fullDfKey, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/20250613_FullDfData.csv")

##dfgenes = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/GenesInDeficiency.csv"))

##fulldf = cbind(dfgenes, dfdata[dfgenes$FBab,])
##row.names(fulldf) = 1:nrow(fulldf)
##fulldf$Modifier[fulldf$Modifier ==  "Supressor"] = "Suppressor"
##fulldf$Modifier[fulldf$Modifier ==  "No effect"] = "No Effect"
##new=fulldf

##olddf = DfKey$Deficiency
##newdf = unique(new$Deficiency)
##oldfb = old$FBgn
##newfb = unique(new$FBgn)
##oldsym = old$Symbol
##newsym = unique(new$Symbol)

fullDfKey = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/20250613_FullDfData.csv"),row.names = 1)

row.names(fullDfKey) = fullDfKey$FBab

#Build a fresh gene id data table
gene.chr = as.data.frame(org.Dm.egCHR)
row.names(gene.chr) = gene.chr$gene_id
gene.start = as.list(org.Dm.egCHRLOC)
gene.start = lapply(gene.start, min)
gene.start = unlist(gene.start)
gene.end = as.list(org.Dm.egCHRLOCEND)
gene.end = lapply(gene.end, max)
gene.end = unlist(gene.end)
FBid = as.data.frame(org.Dm.egFLYBASE)
row.names(FBid) = FBid$gene_id
Symbols = as.data.frame(org.Dm.egSYMBOL)
row.names(Symbols) = Symbols$gene_id

gene.location = data.frame(FBgn = FBid[names(gene.start), "flybase_id"],
                           Symbol = Symbols[names(gene.start), "symbol"],
                           ensembl = names(gene.start),
                           CHR = gene.chr[names(gene.start), "chromosome"],
                           Start = abs(gene.start),
                           End = abs(gene.end[names(gene.start)]),
                           Strand = gene.start/abs(gene.start))

gl = apply(gene.location[,c("Start", "End")], MARGIN = 1, min)
gr = apply(gene.location[,c("Start", "End")], MARGIN = 1, max)
Dfl = apply(fullDfKey[,c("Start", "End")], MARGIN = 1, min)
Dfr = apply(fullDfKey[,c("Start", "End")], MARGIN = 1, max)

genesindf = list()
i=1
while(i <= nrow(fullDfKey)){
  temp = na.omit(gene.location[((gl <= Dfl[i] & gr >= Dfr[i]) |
                               (gl >= Dfl[i] & gl <= Dfr[i]) |
                               (gr >= Dfl[i] & gr <= Dfr[i])) &
                               gene.location$CHR == fullDfKey$Chr[i],])

  genesindf[[fullDfKey$Deficiency[i]]] = temp$FBgn
  i=i+1
}
